Scrutinizing the linear variable importance of local explanations of non-linear models with animated linear projections

### 2022-01-02 _paper.rmd
Couple laste touch up, and linting while refreshing thesis_ns

<!-- Interpretability important for fighting baises -->
Interpretability is vital for exploring and protecting against potential biases in any model. For instance, models regularly pick up on biases in the training data that have observed influence on the response variable, which is then built into the model. Such biases include sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018]. Variable-level interpretability of models is essential in the evaluation models for such biases.

<!-- Interpretability for data drift -->
Another concern is that of data drift. Data drift is an unexpected shift in support of the variables. Non-linear models are regularly more sensitive to these changes. The interpretability of the model means more transparency to see where models' predictions may be plausible or completely unreliable.

@@

<!-- Reactive vs exploration interactions -->
The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation. These reactive inputs make the software flexible and extensible. The application also has more exploratory interactions to help link points across displays and reveal structure found in differnt spaces.

@@

```{r interactions, echo=F, out.width = "100%", fig.cap = "Illustration of data explorations interactions in the global view. This view has linked brushing, where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below. Tooltips display when hovering over an observation."}
knitr::include_graphics("./figures/app_interactions.PNG")
```

@@

It is vital to mitigate the render time of visuals, especially when users may want to iterate many times. All computational operations should be prepared before runtime. The work remaining when an appication is ran is solely reacting to inputs and rendering of visuals and tables. Below we discuss the steps and details of the reprocessing.

@@

(ref:casepenguins) Species classification of Palmer penguin data. We select a chinstrap penguin that is mislabeled as an adelie. By varying the contribution to bill length, we observe the explanation does not hold when bill length has a significant contribution. The .mp4 animation of this tour can be found at [github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4)
```{r casepenguins, echo=F, out.width = "100%", fig.cap = "(ref:casepenguins)"}
knitr::include_graphics("./figures/case_penguins.png")
```

@@

(ref:casechocolates) Chocolates data type classification (milk or dark). We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate in light of the values on the nutritional label. We vary the contribution of calories from fat. Animated tour can be found at [github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
```{r casechocolates, echo=F, out.width = "100%", fig.cap = "(ref:casechocolates)"}
knitr::include_graphics("./figures/case_chocolates.png")
```

@@

(ref:caseames) Ames housing 2018 regressing log sales price [2018 USD]. Because the SHAP values are relatively well distributed across the variables, there is a lot of redundant information to explain the gap between select observations. This case may not be ideal for the cheem methodology, though we did find a near singular frame in the manual tour not, something we came looking for, but certainly, an exciting feature to find. The corresponding animation is at [github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4)
```{r caseames, echo=F, out.width = "100%", fig.cap = "(ref:caseames)"}
knitr::include_graphics("./figures/case_ames2018.png")
```

@@

# Discussion {#sec:discussion}

@@

We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader $\text{MI}^\text2$ lab group for the __DALEX__ ecosystem of __R__ and __Python__ packages. This research was supported by Australian Government Research Training Program (RTP) scholarships. Thanks to Jieyang Chong for helping proofread this piece.

>>>>

<!-- Interpretability & baises -->
Interpretability is vital for exploring and protecting against potential biases in any model. For instance, models regularly pick up on biases in the training data that have observed influence on the response variable, which is then built into the model. Such biases include sex [@dastin_amazon_2018; @duffy_apple_2019], race [@larson_how_2016], and age [@diaz_addressing_2018]. Variable-level interpretability of models is essential in the evaluation models for such biases.

<!-- Interpretability & data drift -->
Another concern is that of data drift. Data drift is an unexpected shift in support of the variables. Non-linear models are regularly more sensitive to these changes. The interpretability of the model means more transparency to see where models' predictions may be plausible or completely unreliable.

@@

<!-- Reactive vs exploration interactions -->
The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation. These reactive inputs make the software flexible and extensible. The application also has more exploratory interactions to help link points across displays and reveal structure found in different spaces.

@@

```{r interactions, echo=F, out.width="100%", fig.cap = "Illustration of data explorations interactions in the global view. This view has linked brushing, where observations selected in one facet are highlighted in the other facets and populate an interactive tabular display below. Tooltips display when hovering over an observation."}
knitr::include_graphics("./figures/app_interactions.PNG")
```

@@

It is vital to mitigate the render time of visuals, especially when users may want to iterate many times. All computational operations should be prepared before runtime. The work remaining when an application is ran is solely reacting to inputs and rendering of visuals and tables. Below we discuss the steps and details of the reprocessing.

@@

(ref:casepenguins) Species classification of Palmer penguin data. We select a chinstrap penguin that is mislabeled as an adelie. By varying the contribution to bill length, we observe the explanation does not hold when bill length has a significant contribution. The .mp4 animation of this tour can be found at [github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_penguins.mp4)
```{r casepenguins, echo=F, out.width="100%", fig.cap = "(ref:casepenguins)"}
knitr::include_graphics("./figures/case_penguins.png")
```

@@

(ref:casechocolates) Chocolates data type classification (milk or dark). We select a chocolate labeled as dark though a random forest model predicts it to be milk chocolate in light of the values on the nutritional label. We vary the contribution of calories from fat. Animated tour can be found at [github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_chocolates.mp4).
```{r casechocolates, echo=F, out.width="100%", fig.cap = "(ref:casechocolates)"}
knitr::include_graphics("./figures/case_chocolates.png")
```

@@

(ref:caseames) Ames housing 2018 regressing log sales price [2018 USD]. Because the SHAP values are relatively well distributed across the variables, there is a lot of redundant information to explain the gap between select observations. This case may not be ideal for the cheem methodology, though we did find a near singular frame in the manual tour not, something we came looking for, but certainly, an exciting feature to find. The corresponding animation is at [github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4](https://github.com/nspyrison/cheem_paper/blob/main/figures/case_ames2018.mp4)
```{r caseames, echo=F, out.width="100%", fig.cap = "(ref:caseames)"}
knitr::include_graphics("./figures/case_ames2018.png")
```

@@

# Discussion {#sec:cheemdiscussion}

@@

We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader $\text{MI}^\text2$ lab group for the __DALEX__ ecosystem of __R__ and __Python__ packages. This research was supported by Australian Government Research Training Program (RTP) scholarships. Thanks to Jieyang Chong for helping proofread this article.

### 2022-01-02 _paper.tex
Couple laste touch up, and linting while refreshing thesis_ns

The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation. These reactive inputs make the software flexible and extensible. The application also has more exploratory interactions to help link points across displays and reveal structure found in differnt spaces.

@@

\hypertarget{sec:discussion}{%
\section{Discussion}\label{sec:discussion}}

@@

We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader \(\text{MI}^\text2\) lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages. This research was supported by Australian Government Research Training Program (RTP) scholarships. Thanks to Jieyang Chong for helping proofread this piece.

>>>>

The application has several reactive inputs that affect the data used, aesthetic display, and tour manipulation. These reactive inputs make the software flexible and extensible. The application also has more exploratory interactions to help link points across displays and reveal structure found in different spaces.

@@

\hypertarget{sec:cheemdiscussion}{%
\section{Discussion}\label{sec:cheemdiscussion}}

@@

We would like to thank Professor Przemyslaw Biecek for his input early in the project and to the broader \(\text{MI}^\text2\) lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Python} packages. This research was supported by Australian Government Research Training Program (RTP) scholarships. Thanks to Jieyang Chong for helping proofread this article.
